### 2020-02-11 Paper/paper.tex
minor changes

%----------------------------------------------------------------------------------------
%	HEADER
%----------------------------------------------------------------------------------------
% !TeX spellcheck = en_US

\documentclass[12pt, a4paper, oneside]{article}

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage{breqn} % for line breaks in equations

\usepackage[top=3cm, left=3cm, right=2cm, bottom=2.5cm]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage[onehalfspacing]{setspace}

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

% TODO do we need a abstract?
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Water Management $\bullet$ Seminar Computational Economics $\bullet$ Winter Semester 2019/2020} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\setlength{\parindent}{0em}

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{graphicx} % for figures

\usepackage{longtable}

% packages used for citations
\usepackage[
backend=bibtex,
style=alphabetic,
citestyle=authoryear,
natbib=true
]{biblatex}
\addbibresource{collection.bib}

% TODO delete the lorem ipsum
\usepackage{blindtext} % Package to generate dummy text throughout this template

\title{Water Management}
%----------------------------------------------------------------------------------------

\begin{document}

	%----------------------------------------------------------------------------------------
	%	COVER SHEET
	%----------------------------------------------------------------------------------------

	\begin{titlepage}
		\begin{center}
			\vspace*{1cm}

			\textbf{Seminar \textit{Computational Economics}}

			\vspace{0.5cm}
			{\makeatletter{\@title}\makeatother}

			\vspace{1.5cm}

			\textbf{Felix Fink} \\
			stud. 2nd semester M.Sc. Business Administration and Engineering: Mechanical Engineering \\
			Matr. Nr. 345971 \\
			\href{mailto:felix.fink@rwth-aachen.de}{felix.fink@rwth-aachen.de}

			\vspace{1cm}

			\textbf{Niklas Sayer} \\
			stud. 2nd semester M.Sc. Business Administration and Engineering: Materials- and Process Engineering\\
			Matr. Nr. 332404 \\
			\href{mailto:niklas.sayer@rwth-aachen.de}{niklas.sayer@rwth-aachen.de}

			\vspace{1cm}

			\textbf{Martin Wicke} \\
			stud. 4th semester M.Sc. Business Administration and Engineering: Electrical Power Engineering\\
			Matr. Nr. 336301 \\
			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke@rwth-aachen.de}

			\vfill

			Supervisor\\
			Thomas Gehrmann, M.Sc.

			\vspace{0.8cm}

			%\includegraphics[width=0.4\textwidth]{university}

			Computational Economics\\
			RWTH Aachen University\\
			Winter Semester 2019/2020

		\end{center}
	\end{titlepage}

	%----------------------------------------------------------------------------------------
	%	TABLE OF CONTENTS
	%----------------------------------------------------------------------------------------

	\tableofcontents
	\clearpage

	%----------------------------------------------------------------------------------------
	%	ARTICLE CONTENTS
	%----------------------------------------------------------------------------------------

	\section{Introduction}

	\subsection{Motivation}
With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people will have to be fed, which will increase the demand of water for agriculture uses \citep{un2019}.
Already today, around one third of the world's population live under physical water scarcity \citep{vorosmarty2000global, alcamo2003global, oki2006global}.
With climate change proceeding, rainfall as the natural influx of lakes will become more volatile in some regions while prolonged dry periods will make water an even more precious resource in other areas \citep{guhathakurta2011impact}.
This situation will get even worse due to the projected increase in the demand of water per capita in many regions of the world, driven by lifestyle factors, especially diet \citep{vorosmarty2000global}.
Despite all those challenges, research suggests that just by using water more efficiently, we will be able to meet the acute water, environment and poverty challenges facing us over the next 50 years \citep{molden2013water}.
Therefore, the optimal allocation and sustainable use of the available water will be of increasing importance in the future.

\subsection{Assumptions and Conditions}

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of all different parties have to be considered.
The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities.
Note that while looking at just two agents is a vast simplification compared to a real allocation problem, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in figure ~\ref{fig:Sequence}.
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

\begin{equation}
	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
\end{equation}

where $x$ represents the amount of water to be used for irrigation in a year. The preference parameters are $a_1 = 1$ and $b_1 = -2$.\\\\

During the summer months, citizens benefit from the remaining water due to recreational use (Step 2).
This can be described by the recreational users' utility function which is

\begin{equation}
  G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
\end{equation}

with the additional paramter $s$ that indicates the water level at the beginning of the year.
For recreational users, the preference parameters shell be $a_2 = 2$ and $b_2 = -3$.\\\\

At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly following a lognormal distribution \citep{oosterbaan1994frequency}:
\begin{equation}
	r \sim lognormal(\mu, \sigma^2)
\end{equation}

with parameters $\mu = 0$ and $\sigma^2 = 1$.\\\\

Other parameters are the maximum capacity of the reservoir
\begin{equation}
	M = 7
\end{equation}

and the discount factor for future utility

\begin{equation}
	\beta = 0.9
\end{equation}

\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/CESchaubild_Cut.png}
	\caption{Sequence of Events in One Period}
	\label{fig:Sequence}
\end{figure}

Using dynamic programming in combination with a stochastic simulation of the rainfall, we determined the optimal irrigation policy that maximizes the combined utility of farmers and citizens for a number of years $T$.
In the main part we will give a brief summary of the underlying theory and the optimization problem.
After that we will present the results and our conclusions for the original model and its modifications, as well as possible shortfalls of our analysis.
At the end we also give a little outlook into further possible extensions.

\subsubsection{Modification 1: Variation of the Discount Factor}
We also study the effects of different betas on the optimal irrigation policy. 
The discount factor takes into account that agents in a simple two-stage model often value current consumption over future consumption.
It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them easily comparable \citep{kruschwitz2014investitionsrechnung}.
We take a look at the discount factors x and x.

\subsubsection{Modification 2: Increased Rainfall Volatility}
To consider the predicted impacts of climate change \citep{guhathakurta2011impact}, we look at the effect of increased rainfall volatility. 
We do this by using a  $\sigma^2$ of x.

\subsubsection{Modification 3: Using Real Data for Rainfall}
Instead of simulating the rainfall with random numbers, we also use real data.
Comparing the results then allows us to draw conclusions on the accuracy of the simulation for that region. 
We use a dataset aquired from \cite{kaggle:2019}, that contains real rainfall data for the last 65 years in different regions of Bangladesh. % Zitation anpassen

\subsubsection{Modification 4: Taking Evaporation into Account}
The evaporation of water from a water reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}.
It therefore should definitely be considered in a real model that is used to determine the optimal irrigation strategy.
The dataset that we use for the real rainfall also contains many other weather parameters.
With the mean temperature $T$, the elevation $h$, the latitude $A$ and the dew point $T_d$ we calculate an estimate for evaporation of water in mm per day for that specific location, using the formula \citep{linacre1977simple}

\begin{equation}
	{E_0} = \frac{700 * (T + 0.006h) / (100 - A) + 15 * (T - T_d)}{(80 - T)}
\end{equation}

and then multiply the result with 365.25 to get the estimated evaporation in mm per year. 
Note, however, that due to a lack of data on the real mean temperature in a year, we use the average of the minimum and maximum temperature of every month for all months in a year as $T$. \\
The average dew point $T_d$ can be estimated by

\begin{equation}
	{T_d} = T - (\frac{100 - RH}{5})
\end{equation}

where RH is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\

Note that all modifications, except Modification 4, are based on the original model without any modifications.
Modification 4 is an extension of Modification 3.

	%------------------------------------------------

	\section{Main Part}
	\subsection{Theory}

	In this paper, the theoretical knowledge obtained in the seminar shall be applied.

	\subsubsection{Quadrature}
	In many computational applications, formulas 	are used to calculate values of any kind. For example, let the flow of water through a pipe be

	\begin{equation}
	f(t) = sin(\pi t), \; t \in [0,1]
	\end{equation}

Now, it's trivial to compute the flow at any given time using code, simply by implementing the formula as given above. 
However, if we want to compute the amount of water flowing through the pipe in the time interval $[0.25, 0.5]$, we can not calculate that directly, as we would have to know the antiderivative of $f(t)$. 
Apart from this example, it's not always trivial to calculate the antiderivative of a function analytically. 
[TODO: zwei mal however]However it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.

A naive implementation of a quadrature could be to calculate the total area of equidistant rectangles whose height corresponds to the function value of $f$:

	\begin{figure}[ht] 
		\includegraphics[width=1\textwidth]{figures/quadratureA.eps}
		\caption{Quadrature using the Midpoint Rule}
		\label{fig:quadrature-a}
	\end{figure}
This implementation of a quadrature is called \emph{Midpoint Rule}. 
\\

A more sophisticated way to determine the area of an unbounded integral is to use the Gauss-Hermite Formula

	\begin{equation}
	\int_{-\infty}^{\infty} f(x)e^{-x^2} dx \approx \sum_{i=1}^{n}{w_if(x_i)}
	\end{equation}

where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the nth Hermite polynomial. (\cite{seminar:week6})

\subsection{Dynamic Programming}

A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). 
In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2). 
This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard o the state resulting from the first decisions" (\cite{bellman:1957}, p.2).

\subsection{The Bellman Equation}

The so-called "Bellman Equation" answers the question, assuming a given state, what long-term utility can be expected by taking the best possible action now and at each subsequent step - or simply put: "What is the value $V$ of a state $s$?".

Mathematically, we can express this as  
\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ [TODO: QUOTE Seminar Week 3].

So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$.

\subsection{Implementation} % TODO: Better title?

In a first step, we want to compute the value function $V$ for each state $s$. 
In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
Obviously, the $aux$ matrix can only a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
Thus we discretize both water level $s$ and irrigation amount $x$ as vectors

	\begin{equation}
s = \left( \begin{array}{c}0\\...\\7\end{array}\right),\hspace{2mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
\end{equation}

Note that the maximum reservoir capacity is $7$ and $|s|=|x| = n = 1000$.

For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix

\begin{dmath}
aux = \left[\begin{array}{ccccc} 
u(s_1, x_1)& \emptyset & \emptyset & ... & \emptyset \\ u(s_2, x_1)& u(s_2, x_2)& \emptyset & ... & \emptyset \\                                               u(s_3, x_1)& ... & u(s_3, x_3)& ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              u(s_n, x_1)& ... & ... & ... & u(s_n, x_n)\\                                               \end{array}\right]+\beta*\left[ \begin{array}{ccccc}
V'(s_1') & \emptyset & \emptyset & ... & \emptyset \\                                               V'(s_1') & V'(s_2') & \emptyset & ... & \emptyset \\                                               V'(s_1') & ... & V'(s_3') & ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              V'(s_1') & ... & ... & ... & V'(s_n') \\                                               \end{array}\right]
\end{dmath}

where the upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.

Here, $V'$ represents the value for $V$ in the previous iteration and will be set to $V'(s'_1) = (0,...0)^T$. Iteratively, we use the bellman equation to calculate a new optimal V:

\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

which equals to finding the column $aux$ that maximizes the expected value over all states (TODO: kann man das so sagen?).
We repeat this step over and over again until the our value function $V$ converges (technically: the norm of the difference between $V'$ and $V$ falls below a certain tolerance value). \\

Finally, we want to compute the value of the steady state by using a monte carlo simulation.

	\subsection{bla}
	\begin{figure}[ht] % TODO remove example
		\includegraphics[width=1\textwidth]{figures/example.png}
		\caption{Example Image}
		\label{fig:example-pic}
	\end{figure}
	\blindtext
	\begin{figure}[ht] % TODO remove example
		\begin{longtable}{|p{2,5cm}|p{2,5cm}|p{2,5cm}|p{3cm}|p{3cm}|}
			\hline
			&\multicolumn{4}{c|}{\textbf{Klasse}} \\
			\hline
			& \textbf{S0} & \textbf{S1} & \textbf{S2} & \textbf{S3} \\
			\hline
			\textbf{lorem} & ipsum & ipsum & ipsum & ipsum \\
			\hline
		\end{longtable}
		\label{fig:example-figure}
		\caption{Example Table}
	\end{figure}
	\blindtext
	\blindtext
	\blindtext

	%------------------------------------------------
	\newpage
	\section{Results}
	This chapter will discuss the outcomes and results of the implemented functionality in MATLAB.
	The discussion will be started on the basic questions asked by the project description.
	Furthermore we made several modifications to the task which will be discussed additionally.
	All results will be visualized by plots made within our simulation in MATLAB.
	\subsection{Optimal irrigation policy}
	When questioning what the optimal irrigation policy is, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like.
	The value function as posed in % TODO put links to theory here
	contains the maximum possible usage that can be gained by irrigating a specific amount of water under a given water level. 
	Figure \ref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of 0 and 7. 
	\begin{figure}[ht]
		\includegraphics[width=1\textwidth]{figures/value_function.eps}
		\caption{Value Function}
		\label{fig:value-function}
	\end{figure}
	\newline
	Every value along the possible water levels leads to an irrigation amount that optimizes the value for all parties in the given situation. Summarizing these values over the range of possible water levels results in a optimal irrigation policy which is shown in figure \ref{fig:optimal-irrigation-policy}.
	The plot shows a rising proposed irrigation amount with rising water level. There are 2 particularities that can be found within the graph:
	\begin{itemize}
		\item Due to the effect that a almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
		\item The diminishing margin utility of farmers lead to a decreasing gradient for the proposed irrigation amount with increasing water level.
	\end{itemize}
	\begin{figure}[ht]
		\includegraphics[width=1\textwidth]{figures/OptimalIrrigationPolicy.eps}
		\caption{Optimal Irrigation Policy}
		\label{fig:optimal-irrigation-policy}
	\end{figure}
	\subsection{Steady State}
	% TODO maybe move to theory?
	In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \cite{gagniuc2017markov}. 
	The variable defining the behavior of our system is the water level. 
	Because of this, we have to calculate the difference in water level for each period. 
	Due to the fact, that our water level highly depends on the randomly distributed rainfalls that fill the reservoir at end of each period, it is not possible to find a state steady. % TODO: auch noch evaporation nennen?
	To accomplish this, we defined our steady state, to be a state, where the mean water level of the last 200 periods did not change more than $1x10^{-7}$ from a period to the next.
	A exemplary graph of the mean in every period and the difference to its predecessor is shown in figure \ref{fig:steadyState}.
	Once a period is found, where the mean does not change more than the given value, the mean of that period is defined as steady state water level.
	The water level found to be a steady state and the period in which this condition occurs changes from calculation to calculation due to the fact, that the rain fall is still a random distribution.
	\begin{figure}[ht]
		\includegraphics[width=1\textwidth]{figures/steadyState.eps}
		\caption{Exemplary Calculation of the Steady State}
		\label{fig:steadyState}
	\end{figure}
	\subsection{Steady State Probability}
	As mentioned above, the steady state changes from simulation to simulation.
	In order to find out, how the steady states are distributed and what the probability of their occurrences are, we did a Monte Carlo simulation.
	To do so, we ran our code 1000 times and put the resulting water levels into bins with the size of 0.1 each.
	We then counted the occurrences, normalized them and calculated the corresponding probability.
	The result is put in a histogram and shown in figure \ref{fig:steadyStateMonteCarlo}
	\begin{figure}[ht]
		\includegraphics[width=1\textwidth]{figures/steadyStateMonteCarlo.eps}
		\caption{Steady State Probability}
		\label{fig:steadyStateMonteCarlo}
	\end{figure}
	\subsection{Effect of Changed Discount Factors to Optimal Irrigation Policy}
	\subsection{Changed Volatility of Rainfalls}

	\printbibliography

	\newpage

	%----------------------------------------------------------------------------------------
	%	STATUTORY DECLARATION IN LIEU OF AN OATH
	%----------------------------------------------------------------------------------------

	\begingroup
	\begin{center}
		\large
		\textbf{Eidesstattliche Versicherung}
	\end{center}

	\vspace{0.3cm}

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
		\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
	\end{tabular}
	\vspace{0.3cm}

	Ich versichere hiermit an Eides Statt, dass ich die vorliegende Arbeit mit dem Titel

	\begin{center}
		{\makeatletter{\emph{{\@title}}}\makeatother}
	\end{center}

	selbständig und ohne unzulässige fremde Hilfe erbracht habe. Ich habe keine anderen als die angegebenen Quellen und Hilfsmittel benutzt. Für den Fall, dass die Arbeit zusätzlich auf einem Datenträger eingereicht wird, erkläre ich, dass die schriftliche und die elektronische Form vollständig übereinstimmen. Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen.

	\vspace{0.6cm}

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
		Ort, Datum & Unterschrift \\
	\end{tabular}

	\vspace{0.6cm}

	\begin{small}
		\textbf{Belehrung:}\\
		Wer vorsätzlich gegen eine die Täuschung über Prüfungsleistungen betreffende Regelung einer Hochschulprüfungsordnung verstößt, handelt ordnungswidrig. Die Ordnungswidrigkeit kann mit einer Geldbuße bis zu 50 000 Euro geahndet werden. Zuständige Verwaltungsbehörde für die Verfolgung und Ahndung von Ordnungswidrigkeiten ist die Kanzlerin oder der Kanzler der RWTH Aachen. Im Falle eines mehrfachen oder sonstigen schwerwiegenden Täuschungsversuches kann der Prüfling zudem exmatrikuliert werden (§ 63 Abs. 5 HG NRW).

		\textbf{§ 156 StGB: Falsche Versicherung an Eides Statt}\\
		Wer vor einer zur Abnahme einer Versicherung an Eides Statt zuständigen Behörde eine solche Versicherung falsch abgibt oder unter Berufung auf eine solche Versicherung falsch aussagt, wird mit Freiheitsstrafe bis zu drei Jahren oder mit Geldstrafe bestraft.

		\textbf{§ 161 StGB: Fahrlässiger Falscheid; fahrlässige falsche Versicherung an Eides Statt}\\
		(1) Wenn eine der in den §§ 154 bis 156 bezeichneten Handlungen aus Fahrlässigkeit begangen worden ist, so tritt Freiheitsstrafe bis zu einem Jahr oder Geldstrafe ein.\\
		(2) Straflosigkeit tritt ein, wenn der Täter die falsche Angabe rechtzeitig berichtigt. Die Vorschriften des § 158 Abs. 2 und 3 gelten entsprechend
	\end{small}

	\vspace{0.6cm}

	Die vorstehende Belehrung habe ich zur Kenntnis genommen:

	\vspace{0.6cm}

	\begin{tabular}{@{}p{8cm}p{5.8cm}}
		\underline{\smash{Aachen, den \today}} & \underline{\hspace{5.8cm}}\\
		Ort, Datum & Unterschrift \\
	\end{tabular}

	\endgroup
	\clearpage

\end{document}