### 2020-02-24 Paper/paper.tex
changes in chapters 1 and 3 (3.1-3.5)

%----------------------------------------------------------------------------------------
%	HEADER
%----------------------------------------------------------------------------------------
% !TeX spellcheck = en_US

\documentclass[12pt, a4paper, oneside]{article}

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage{breqn} % for line breaks in equations

\usepackage[top=3cm, left=3cm, right=2cm, bottom=2.5cm]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage[onehalfspacing]{setspace}

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles
\newcommand{\sectionbreak}{\clearpage}

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Water Management $\bullet$ Seminar Computational Economics $\bullet$ Winter Semester 2019/2020} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\setlength{\parindent}{0em}

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{graphicx} % for figures
\usepackage{subcaption}
\usepackage{wrapfig}

\usepackage[noabbrev]{cleveref}

\usepackage{longtable}

% packages used for citations
\usepackage[
backend=bibtex,
style=ieee,
citestyle=authoryear,
natbib=true
]{biblatex}
\addbibresource{collection.bib}

\title{Water Management}
%----------------------------------------------------------------------------------------

\begin{document}
	\let\origref\ref % The original \ref command is now accessible by \origref
	\let\ref\cref    % The command \ref uses \cref from the cleveref package now

	%----------------------------------------------------------------------------------------
	%	COVER SHEET
	%----------------------------------------------------------------------------------------

	\begin{titlepage}
		\begin{center}
			\vspace*{1cm}

			\textbf{Seminar \textit{Computational Economics}}

			\vspace{0.5cm}
			{\makeatletter{\@title}\makeatother}

			\vspace{1.5cm}

			\textbf{Felix Fink} \\
			stud. 2nd semester M.Sc. Business Administration and Engineering:\\Mechanical Engineering \\
			Matr. Nr. 345971 \\
			\href{mailto:felix.fink@rwth-aachen.de}{felix.fink@rwth-aachen.de}

			\vspace{1cm}

			\textbf{Niklas Sayer} \\
			stud. 2nd semester M.Sc. Business Administration and Engineering:\\Materials- and Process Engineering\\
			Matr. Nr. 332404 \\
			\href{mailto:niklas.sayer@rwth-aachen.de}{niklas.sayer@rwth-aachen.de}

			\vspace{1cm}

			\textbf{Martin Wicke} \\
			stud. 4th semester M.Sc. Business Administration and Engineering:\\Electrical Power Engineering\\
			Matr. Nr. 336301 \\
			\href{mailto:martin.wicke@rwth-aachen.de}{martin.wicke@rwth-aachen.de}

			\vfill

			Supervisor\\
			Thomas Gehrmann, M.Sc.

			\vspace{0.8cm}

			%\includegraphics[width=0.4\textwidth]{university}

			Computational Economics\\
			RWTH Aachen University\\
			Winter Semester 2019/2020

		\end{center}
	\end{titlepage}

	%----------------------------------------------------------------------------------------
	%	TABLE OF CONTENTS
	%----------------------------------------------------------------------------------------

	\tableofcontents
	\clearpage

	%----------------------------------------------------------------------------------------
	%	ARTICLE CONTENTS
	%----------------------------------------------------------------------------------------

	\section{Introduction}

	\subsection{Motivation}
With world population expected to hit 9.7 billion in 2050 and 11.2 billion in 2100, more and more people will have to be fed, which will increase the demand of water for agriculture uses \citep{un2019}.
Already today, around one third of the world's population live under physical water scarcity \citep{vorosmarty2000global, alcamo2003global, oki2006global}.
With climate change proceeding, rainfall as the natural influx of lakes will become more volatile in some regions while prolonged dry periods will make water an even more precious resource in other areas \citep{guhathakurta2011impact}.
This situation will get even worse due to the projected increase in the demand of water per capita in many regions of the world, driven by lifestyle factors, especially diet \citep{vorosmarty2000global}.
However, research suggests that just by using water more efficiently, we will be able to meet the challenges in water, environment and poverty facing us over the next 50 years \citep{molden2013water}.
Therefore, the optimal allocation and sustainable use of the available water will be of increasing importance in the future.

\subsection{Assumptions and Conditions}

In this work, we want to examine the variations in use of water from a (hypothetical) water reservoir.
One of the most important uses of surface water is the irrigation of fields for food production. However, as farming and food production are private enterprises and lakes are public goods, interests of different parties have to be considered. 
The water reservoir to be considered has two stakeholders: farmers, who want to use the water for irrigation purposes and citizens, who want to use the lake for recreational activities. 
Though looking at just two agents is a vast simplification compared to real allocation problems, one can easily add more stakeholders to the model, as long as there is enough computational power.
The sequence of events for every period is displayed in \ref{fig:Sequence}. %TODO Die Figure ist im PDF zu weit weg gerutscht! 
At the beginning of every year (Step 1) we have to decide how much of the water from the reservoir is used for the irrigation of fields, which benefits the farmers' harvest.\\\\
The farmers utility function follows the form

\begin{equation}
	F(x) = \frac{a_1}{1+b_1} * x^{1+b_1}
\end{equation}

where $x$ represents the amount of water to be used for irrigation during a period. 
The preference parameters are given as $a_1 = 1$ and $b_1 = -2$.\\\\

During the summer months, citizens benefit from the remaining water in their recreational activities (Step 2).
This can be described by the recreational users' utility function

\begin{equation}
  G(s, x) = \frac{a_2}{1+b_2} * (s-x)^{1+b_2}
\end{equation}

with the additional parameter $s$ that indicates the water level at the beginning of the year.
For recreational users, the preference parameters shall be $a_2 = 2$ and $b_2 = -3$.\\\\

At the end of the year, water in the reservoir is replenished by rainfall (Step 3), which we can assume to be randomly distributed following a lognormal distribution \citep{oosterbaan1994frequency}: %TODO cite steht über den Rand hinaus
\begin{equation}
	r \sim lognormal(\mu, \sigma^2)
\end{equation}

with parameters $\mu = 0$ and $\sigma^2 = 1$.\\\\

Other parameters are the maximum capacity of the reservoir
\begin{equation}
	M = 7
\end{equation}

and the discount factor for future utility

\begin{equation}
	\beta = 0.9
\end{equation}

\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/CESchaubild_Cut.png}
	\caption{Sequence of Events in One Period}
	\label{fig:Sequence}
\end{figure}

In this work, we will determine an optimal irrigation policy that maximizes the combined utility of farmers and citizens over $T$ years by using dynamic programming in combination with a stochastic simulation of the rainfall. 
In the main part we will give a brief summary of the underlying theory and the optimization problem. 
After that we will present the results and our conclusions for the original model and its modifications, as well as possible shortfalls of our analysis. %TODO statt "possible shortfalls" besser "possible enhancements" schreiben? (klingt positiver?)
Finally, we will give an outlook for further possible extensions.

\subsubsection{Modification 1: Variation of the Discount Factor}
In this modification, we're interested in studying the effects of different betas on the optimal irrigation policy. 
The discount factor takes into account that agents in a simple two-stage model often value current consumption over future consumption.
It expresses how much one unit of future utility is worth in terms of current utility and therefore makes them mathematically comparable \citep{kruschwitz2014investitionsrechnung}. %TODO wirklich deutsche Citation verwenden? 
We take a look at the discount factors from X to Y. %TODO tatsächliche Grenzen einfügen

\subsubsection{Modification 2: Increasing the Volatility of Rainfalls}
In order to take into account expected future impacts of climate change \citep{guhathakurta2011impact}, we consider the effects of increased volatility of precipitation. %TODO cite steht über den Rand hinaus
We do so by setting $\sigma$ to 2. %TODO Das ist noch nicht ganz richtig

\subsubsection{Modification 3: Using Real Data for Rainfalls}
Instead of simulating the rainfall with randomly generated numbers, we want to use real data. 
We make use of a dataset aquired from \citep{kaggle:2019}, which contains rainfall and other weather related data from the last 65 years in different regions of Bangladesh (here we will focus on the city Barisal). %TODO Zitation anpassen
In order to make the dataset compatible to the basic model, we assume an exemplary surface-to-volume-ratio of the reservoir. For the sake of this simulation, we multiply the given rainfall by a factor of $4$, in order to appropriately replenish the reservoir.

\subsubsection{Modification 4: Taking Evaporation into Account} \label{sec:intro-mod4}
The evaporation of water from a reservoir is the most important natural "outflow" to be considered when it comes to managing water levels \citep{tanny2008evaporation}. 
To determine a more realistic optimal irrigation strategy, we should therefore consider evaporation effects in our model. 
The dataset that we use for the real rainfall also contains many other weather parameters.
The dataset from \citep{kaggle:2019} does not contain evaporation information directly. However, with the mean temperature $T$ in °C, the elevation $h$ in m, the latitude $A$ in °S and the dew point $T_d$ in °C we can calculate an estimate for the evaporation of water in mm per day for a specific location, using the formula \citep{linacre1977simple}

\begin{equation}
	{E_0} = \frac{700 * (T + 0.006h) / (100 - A) + 15 * (T - T_d)}{(80 - T)}
\end{equation}

and then multiply the result with 365.25 to get the estimated evaporation in mm per year. 
Note, however, that due to a lack of data on the real mean temperature in a year, we use the average of the minimum and maximum temperature of every month for all months in a year as $T$. \\
The average dew point $T_d$ can be estimated by

\begin{equation}
	{T_d} = T - (\frac{100 - RH}{5})
\end{equation}

where $RH$ is the average relative humidity of the year in percent \citep{lawrence2005relationship}.\\\\

%TODO sind diese beiden Sätze so schön? oder nötig?
Note that all modifications, except Modification 4, are based on the original model without any modifications. 
Modification 4 is an extension of Modification 3.

	%------------------------------------------------

	\section{Main Part}
	\subsection{Theory}

	In this paper, the theoretical knowledge obtained in the seminar shall be applied.

	\subsubsection{Quadrature}
	In many computational applications, formulas 	are used to calculate values of any kind. For example, let the flow of water through a pipe be

	\begin{equation}
	f(t) = sin(\pi t), \; t \in [0,1]
	\end{equation}

Now, it's trivial to compute the flow at any given time using code, simply by implementing the formula as given above. 
However, if we want to compute the amount of water flowing through the pipe in the time interval $[0.25, 0.5]$, we can not calculate that directly, as we would have to know the antiderivative of $f(t)$. 
Apart from this example, it's not always trivial to calculate the antiderivative of a function analytically. 
Nonetheless it is possible compute the area below (almost) any function using numerical integration, also called numerical quadrature.

A naive implementation of a quadrature could be to calculate the total area of equidistant rectangles whose height corresponds to the function value of $f$:

	\begin{figure}[ht] 
		\includegraphics[width=1\textwidth]{figures/quadratureA.eps}
		\caption{Quadrature using the Midpoint Rule}
		\label{fig:quadrature-a}
	\end{figure}
This implementation of a quadrature is called \emph{Midpoint Rule}. 
\\

A more sophisticated way to determine the area of an unbounded integral is to use the Gauss-Hermite Formula

	\begin{equation}
	\int_{-\infty}^{\infty} f(x)e^{-x^2} dx \approx \sum_{i=1}^{n}{w_if(x_i)}
	\end{equation}

where the weights $w_i$ are given as the solution to a linear system of equations and the nodes $x_i$ correspond to the zeros of the n-th Hermite polynomial \citep{seminar:week6}.

\subsubsection{Dynamic Programming}

A common technique to solve optimization or combinatorial problems is called Dynamic Programming. 
It is based on the idea of dividing problems into smaller subproblems, solving these recursively and caching intermediate results (\textit{memoization}). % Nicht eher "memoRization"?
In his paper "The Theory Of Dynamic Programming" from 1957, Richard Bellman aimed to provide a solution to multi-stage-decision processes. 
He defines a \textit{policy} as "a sequence of decisions" (\cite{bellman:1957}, p.2), and an \textit{optimal policy} as "a policy, which is most advantageous according to some preassigned criterion" (\cite{bellman:1957}, p.2). 
This is reflected in Bellman's Principle of Optimality: "An optimal policy has the property that whatever initial state and initial decisions are, the remaining decisions must constitute an optimal policy with regard to the state resulting from the first decisions" (\cite{bellman:1957}, p.2).

\subsubsection{The Bellman Equation}

The so-called "Bellman Equation" answers the question, assuming a given state, what long-term utility can be expected by taking the best possible action now and at each subsequent step - or simply put: "What is the value $V$ of a state $s$?".

Mathematically, we can express this as  
\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}

where $u(s,x)$ is the utility to gain from the control variable (action) $x$ in state $s$ \citep{seminar:week3}.

So the value of the control variable to be set is the one that maximizes both the utility of the current state plus (recursively) the value $V$ of the next state $s'$ which is discounted by a factor of $\beta$. 
%TODO hier noch auf die Definition von beta eingehen? Was ist beta und warum ist das relevant?
\subsection{Implementation}
\subsubsection{Computation of the value function}

In a first step, we want to compute the value function $V$ for each state $s$. 
In order to do so, we need to calculate an auxiliary matrix $aux$ that contains the combined utilities of the farmers and the recreational users. 
Obviously, the $aux$ matrix can only contain a finite, discrete amount of values. In this work, a resolution of $n = 1000$ different water levels and irrigation amounts should suffice. 
Thus we discretize both water level $s$ and irrigation amount $x$ as vectors

	\begin{equation}
s = \left( \begin{array}{c}0\\...\\7\end{array}\right),\hspace{2mm}x = \left( \begin{array}{c}0\\...\\7\end{array}\right)
\end{equation}
Note that the maximum reservoir capacity is $7$ and $s, x \in R^{1000}$.

For a combined utility of the farmers and the recreational users  $u(s_i,x_j) = F(x_j) + G(s_i,x_j)$, depending on the water level $s_i$ and the amount of irrigation $x_i$ we get an auxiliary matrix

\begin{dmath}
aux = \left[\begin{array}{ccccc} 
u(s_1, x_1)& \emptyset & \emptyset & ... & \emptyset \\ u(s_2, x_1)& u(s_2, x_2)& \emptyset & ... & \emptyset \\                                               u(s_3, x_1)& ... & u(s_3, x_3)& ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              u(s_n, x_1)& ... & ... & ... & u(s_n, x_n)\\                                               \end{array}\right]+\beta*\left[ \begin{array}{ccccc}
V'(s_1') & \emptyset & \emptyset & ... & \emptyset \\                                               V'(s_1') & V'(s_2') & \emptyset & ... & \emptyset \\                                               V'(s_1') & ... & V'(s_3') & ... & \emptyset \\                                              ... & ... & ... & ... & \emptyset \\                                              V'(s_1') & ... & ... & ... & V'(s_n') \\                                               \end{array}\right]
\end{dmath}

where the upper right triangle of this matrix will be empty, as the amount of water, that can be taken from the reservoir must be smaller or equal to the amount of water in the reservoir.

Here, $V'$ represents the value for $V$ in the previous iteration and will be set to $V'(s'_1) = (0,...0)^T$. Iteratively, we use the bellman equation to calculate a new optimal V:

\begin{equation}
V(s) = \max_{x}(u(s,x) + \beta*V'(s')))
\end{equation}
\label{eq:value-function}

which equals to finding the column $aux$ that maximizes the expected value over all states.
We repeat this step over and over again until the our value function $V$ converges.
In our case the function converges, if the norm of the difference between $V'$ and $V$ falls below a certain tolerance value \\
% TODO Fehlen hier nicht noch ein paar Sätze dazu wie wir die optimal irrigation policy extrahieren?

\subsubsection{Forward Simulation}
In order to simulate the problem, we developed a so called forward simulation which is run once for each period to be simulated.
It consists of several steps that will be described in more detail here.
At the begin of each period the code determines how much water according to the irrigation policy to be used for irrigation.
Since the simulation starts with an empty reservoir the amount irrigated in the first period is zero.
Then the water level for the next iteration is calculated.
This calculation is based on the irrigation amount, the current water level, the amount of rain and the amount that went away by evaporation (see \ref{sec:intro-mod4}).
Afterwards the simulation calculates the difference in water level of the period compared to its predecessor, which will be used for finding a steady state.
% Finally, we want to compute the value of the steady state by using a monte carlo simulation.

%------------------------------------------------
\newpage
\section{Results}
This chapter will discuss the outcomes and results of the implemented functionality in MATLAB.
The discussion will be started on the basic questions asked by the project description.
Furthermore we made several modifications to the task which will be discussed additionally.
All results will be visualized by plots made within our simulation in MATLAB.
\subsection{Optimal irrigation policy}
When examining what an optimal irrigation policy could be, the first question to be asked is how the aggregated value function of both recreational users and farmers looks like. 
The value function as posed in \ref{eq:value-function} contains the maximum possible (expected) utility that can be gained by irrigating a certain amount of water under a given water level in the reservoir. 
Figure \origref{fig:value-function} shows the value function calculated for all possible water levels between the allowed boundaries of $0$ and $7$. 
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/value_function.eps}
	\caption{Value Function}
	\label{fig:value-function}
\end{figure}
\newline
Every value along the possible water levels is mapped to an irrigation amount that optimizes the expected value for all parties in the given situation. 
The resulting optimal irrigation policy is shown in \ref{fig:optimal-irrigation-policy}. 
As the plot shows, a higher water level leads to a higher proposed irrigation amount. There are two particularities that can be found within the graph:
\begin{itemize}
	\item Due to the effect that an almost empty reservoir has a very low use to recreational users, the optimal irrigation strategy suggests a very low irrigation amount whose gradient increases with increasing water level.
	\item As long as the discount factor $\beta$ is greater than $0$, the water within the reservoir has a value in the next period for both recreational users and farmers. Due to this effect, saving waters makes sense and leads to a decreasing gradient for the proposed irrigation amount with increasing water level.
\end{itemize}
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/OptimalIrrigationPolicy.eps}
	\caption{Optimal Irrigation Policy}
	\label{fig:optimal-irrigation-policy}
\end{figure}
\newpage
\subsection{Steady State}
In systems theory, a system or a process is in a steady state, if the variables which define the behavior of the system are unchanging in time \citep{gagniuc2017markov}. 
The variable defining the behavior of our system is the water level. 
Thus, we have to calculate the change in water level for each period. 
Due to the fact that the water level highly depends on the randomly distributed rainfalls (that fill the reservoir at end of each period), it is not possible to find a steady state in which the water level does not change \textit{at all}. 
In this context, we define a steady state to be a state, where the mean water level of the last 200 periods does not change more than $1*10^{-7}$ from one period to the next. %TODO ist dieser Wert noch richtig? Im Code steht im Moment steadyStateTolerance = 0.0001;
A exemplary graph of the mean in every period and the difference to its predecessor is shown in \ref{fig:steadyState}.
Once a period is found where the mean does not change more than the given tolerance, the mean of that period is defined as the steady state of the water level. %TODO hier "moving average" schreiben?
The water level found to be a steady state and the period in which this condition occurs changes from one simulation to the next, since the rainfall is generated randomly following the given distribution.

\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/steadyState.eps}
	\caption{Exemplary Calculation of the Steady State}
	\label{fig:steadyState}
\end{figure}

\subsection{Steady State Probability}
As just mentioned, the steady state changes from simulation to simulation. 
In order to find out how the steady states are distributed and what the probability of their occurrences are, we implement a Monte Carlo simulation. 
To do so, we evaluate the forward simulation 1000 times and put the resulting water levels into bins with the size of 0.1 each. 
We then counted the occurrences, normalized them and calculated the corresponding probability.
The result is put in a histogram and shown in \ref{fig:steadyStateMonteCarlo}.
 %TODO ich (martin) würde vorschlagen, einfach zu schreiben ... "1000 times and calculate a relative frequency distribution" (ist hier erklärt: https://www.sciencedirect.com/topics/mathematics/relative-frequency-distribution)
The highest possibility for a steady state is around 4.2. 
The curve resembles a normal distribution with a standard deviation of $\sigma=0.37$ and peaks at a water level around $4.2$.

\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/steadyStateMonteCarlo.eps}
	\caption{Steady State Probability}
	\label{fig:steadyStateMonteCarlo}
\end{figure}

\subsection{Effect of the Discount Factor on the Optimal Irrigation Policy}
When altering the discount factor $\beta$ in a range from 0 to 1, the optimal irrigation policy changes. %TODO auch hier die tatsächlichen Werte anpassen?!
For small water levels of the reservoir, the optimal irrigation amount is rather insensitive to the discount factor. 
A small $\beta$ ensures that high water levels at the end of one period are only available to a small extent in the next period. 
%TODO Finde die Formulierung nicht so gut: Sie sind ja weiterhin voll verfügbar, aber eben in Zukunft weniger wert. Bzw das könnte man diskutieren. 
This leads to an increasing gap between irrigation amounts over $\beta$ with increasing water level since conserving water does not increase the usage as much. 
One can say that a high $\beta$ prefers a long-term policy while a smaller $\beta$ leads to a rather short-term policy.
%TODO ich (martin) würde diesen absatz so schreiben "A smaller $\beta$ allows more water to be used for irrigation today, optimizing the short-term utility. On the other hand, a higher $\beta$ increases the today's valuation of future utility, and therefore reduces the amount of water to be consumed (irrigated) today. "
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/optimum-policy-over-beta.eps}
	\caption{Optimal irrigation policy as a function of the discount factor $\beta$}
	\label{fig:optimal-irrigation-beta}
\end{figure}
\newpage
\subsection{Increased Volatility of the Rainfalls} %TODO überall use durch utility ersetzen!
More volatile rainfalls means that rainfalls are having a higher standard deviation.
Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed. 
Originally, we assumed $\mu=0$ and $\sigma=1$. Now, we choose $\sigma=2$ and thus have to set $\mu=-1.5$, in order to retain the same mean (see \ref{eq:rainfall}). %TODO ich (martin) hab diesen Absatz so umgeschrieben, einverstanden? Vorher stand hier insbesondere auch (fälschlicherweise?!) $\sigma^2=2$: 
%Since we want to compare two rainfall distributions of the same mean, the median of the distribution has to be changed.
%Using a $\sigma^2=2$ includes setting $\mu=-1.5$ in order for the mean to stay the same.
%(See \ref{eq:rainfall} for reference)

\begin{equation}
	mean = exp(\mu+\frac{\sigma^2}{2}) \label{eq:rainfall}
\end{equation} 
Both distributions are visualized in \ref{fig:rain-distributions}.
\begin{figure}
	\begin{subfigure}{0.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-default.eps}
		\caption{$\mu=0, \sigma=1$}
		\label{fig:rainfall-default}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
		\centering
		\includegraphics[width=1\textwidth]{figures/rainfall-more-volatile.eps}
		\caption{$\mu=-1.5, \sigma=2$}
		\label{fig:rainfall-changed}
	\end{subfigure}
	\caption{Different Rainfall Distributions}
	\label{fig:rain-distributions}
\end{figure}
\newpage
\begin{figure}[ht]
	\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-default.eps}
	\caption{Aggregated Use for lognormal distributed rainfall $(\mu=0, \sigma=1)$}
	\label{fig:aggregated-use-rainfall-default}
\end{figure}
\begin{figure}[ht]
\includegraphics[width=1\textwidth]{figures/aggregated-use-rainfall-more-volatile.eps}
\caption{Aggregated Use for lognormal distributed rainfall $(\mu=-1.5, \sigma=2)$}
\label{fig:aggregated-use-rainfall-more-volatile}
\end{figure}

The following figures demonstrate the effect of the changed volatility. They show the lognormal rainfalls with the original volatility (\ref{fig:aggregated-use-rainfall-default}), and with the increased volatility (\ref{fig:aggregated-use-rainfall-more-volatile}) over 100 periods. 
When comparing both visualizations, we see that more extreme rainfalls significantly affect the water level of the reservoir: there longer periods with a rather low amount of water in the reservoir, but some periods where the reservoir reaches it's capacity limit. 
%When comparing both visualizations,it is noticeable that more volatile rainfall leads to much higher peaks of rain in combination with reduced number of periods with a decent amount of rain. 
%As a consequence the water level in the reservoir fluctuates much more leading to periods with almost no water in the reservoir. 
In the periods with only low precipitation, the farmers use almost no water for irrigation. 
The aggregated utility of farmers and recreational users drops significantly. 
Since the reservoir is capped at an amount of 7, the high peaks of the rainfalls can only partially be used in following (potentially dryer) periods. 
Generally speaking, one can say that a more volatile rain distribution leads to a lower average water level in the reservoir with lower amounts of water to be used for irrigation.
This effect is reflected in the aggregated utility as well. 
The utility of more volatile rainfall is much lower compared to rainfall parameterized with $\mu=0$ and $\sigma=1$. 
Since the mean stays the same in the two cases considered, the value function does not change.

\clearpage
\subsection{Using Real Data for Rainfall}
Figure \origref{fig:rainfall-real-data} shows the distribution of rain from a real data set.
Comparing this with \ref{fig:rainfall-default} shows a similar shape of the distribution. 
Since there are only 65 periods within the real data set is not as well distributed as in the lognormal distributed simulation.
\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/rainfall-real-data.eps}
	\caption{Rainfall Distribution from Real Data}
	\label{fig:rainfall-real-data}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/optimalIrrigationPolicy-real-data.eps}
	\caption{Optimal Irrigation Policy with Real Data}
	\label{fig:optimalIrrigationPolicy-real-data}
\end{subfigure}
\caption{Using Real Data for Rainfall}
\label{fig:real-data}
\end{figure}
In the real data set there are no periods with (almost) no rain while in our simulation there are.
As we can learn from our dataset this assumption is not realistic.
Since the amount of rain is scaled down to match our model, there is no comparison between means of the model possible.
Nevertheless one can say that the overall shape of a lognormal distribution is a sensible estimation for rainfall but for estimating whole years as periods the distribution might to be parameterized differently to fit actual rainfalls. 
As shown in \ref{fig:optimalIrrigationPolicy-real-data} the changed distribution of rainfalls also has an effect on the optimal irrigation policy since it depends on the expected amount of rain.
Due to the higher expected value of rainfalls the optimal irrigation policy is steeper over the water level compared to the simulated rainfall.

\clearpage	
\newpage

\subsection{Taking Evaporation Into Account}
In the following, we want to assess the effect of evaporation on the optimal irrigation policy and the steady state. 
As described in \ref{sec:intro-mod4}, we derive the amount of evaporation over time from other weather related and geographical data. 
The (calculated) evaporation has a low volatility over the years and has value of approximately 1 unit, as \ref{fig:without-vs-with-evaporation} shows. 
Due to the fact that the reservoir has a fixed capacity of 7 units, rainfall that would fill the reservoir above this value is not available for further use. 
When taking evaporation into account, the reservoir never reaches its capacity limits and thus all rainfall will be saved. This leads to an increased volatility of the water level. 
\\

\begin{figure}[h]
\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/results-opt-irrigation-over-time-evap-inactive.eps}
	\caption{Without Evaporation}
	\label{fig:water-level-wo-evap}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/results-opt-irrigation-over-time-evap-active.eps}
	\caption{With Evaporation}
	\label{fig:total-utility-wo-evap}
\end{subfigure}
\caption{Comparison of water level, irrigation amount, rainfall and evaporation over time using real data}
\label{fig:without-vs-with-evaporation}
\end{figure}

As expected, the average water level is lower with evaporation than without, which in return leads to a lower irrigation the farmers are allowed to receive. \\
In the introduction, we stated that due to climate change, prolonged dry periods will occur. We want to reflect this fact in the simulation by scaling the evaporation rate with different factors.

%TODO make the figure better
\begin{figure}[ht]
\begin{subfigure}{0.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/results-evap-factors-water-levels.eps}
	\caption{Water Level}
	\label{fig:water-level-evap}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
	\centering
	\includegraphics[width=1\textwidth]{figures/results-evap-factors-utilities.eps}
	\caption{Total Utility}
	\label{fig:total-utility-evap}
\end{subfigure}
\caption{Water level and log-scaled total utility over time, scaling evaporation with factors $0, 1, 2, 2.3$ (from top to bottom)}
\label{fig:evaporation-factors}
\end{figure}
%\clearpage

\ref{fig:evaporation-factors} shows the total utility and the water level over time, using scaling factors of $0, 1, 2, 2.3$.\footnote{The expected value of the evaporation should not exceed the expected value of the rainfall, as this would make it impossible to calculate the value function. The expected value of the evaporation with a scaling factor of $2.3$ is just below the expected value of the rainfall. } %TODO Fußnote gut so?
We can see that higher amounts of evaporation not only lead to water levels, but also to significantly lower utilities for the recreational users and the farmers. 
Especially when the reservoir completely dries out, the total utility approaches $-\infty$. This happens in periods $7,62,63,64$ and $65$ when applying a factor of $2.3$ to the evaporation.

%TODO improve this figure
\begin{figure}[h]
	\includegraphics[width=1\textwidth]{figures/results-evap-factor-2-3.eps}
	\caption{Evaporation multiplied by a factor of $2.3$ outweighs the rainfall in certain periods} %TODO 
	\label{fig:evap-factor-2.3}
\end{figure}

The reason for this is found in the utility functions of the farmers and the recreational users. Because of $x\leq s$, we get for $s\to 0 \Rightarrow x\to 0$, and as a consequence

\begin{equation}
\lim_{s, x\to0}F(x) = \lim_{s,x\to0}-\frac{1}{x} = -\inf
\end{equation}

as well as

\begin{equation}
\lim_{s, x\to0}G(s,x) = \lim_{s,x\to0}-\frac{1}{(s-x)^2} = -\inf
\end{equation}

%TODO den folgenden Absatz rauswerfen?
How then can these utility functions be interpreted in terms of reality? 
An utility value of negative infinity may be reasonable, if an actor has no alternative actions available. 
In this scenario, we could imagine this to be an isolated village, where the only possibility for the farmers to irrigate their fields is to use water from this particular reservoir, and the only way to have food for survive is to make use of their own fields (nothing else to do, no trading for example), assuming that the farmers want to survive.

\section{Conclusion}
%TODO content
In this work we developed an optimal irrigation strategy for an hypothetical water reservoir, maximizing the sum of utilities for farmers and recreational users.
Using dynamic programming we calculated an aggregated value function based on the given utility functions as described in chapter 2.4.
We applied this value function to a basic model of the water reservoir, by extracting the optimal irrigation path from the value function and simulating water influx in the form of rainfall with a lognormally distributed random variable.
The optimal irrigation policy stipulates quite low irrigation amounts for low water levels, mostly due to the very negative utility citizens receive from an almost empty reservoir (Figure 4).
With $\beta = 0.9$, the optimal irrigation amount tapers off again at the end, due to % TODO: Besprechen! MMn wegen Möglichkeit des Mehrfachnutzens bei Non-irrigation (daher nicht erkennbar bei beta = 0)
In figure 9 you can see one possible scenario in our basic setting with $\mu=0$ and $\sigma=1$.
However, due to the random variable, every forward simulation can look quite different, even though no parameters are changed.
Therefore, the steady state of the water level itself, defined as the moving average of 200 periods (see chapter 3.2), is uncertain.
Running our forward simulation 500 times, we could see that it seems to follow a normal distribution with a mean of around 4.2 and a standard deviation of 0.37. % TODO: StdDev auch im Ergebnisteil erwähenen? 
\\\\
We also looked at how changes in the discount factor $\beta$ influenced the optimal irrigation policy (\ref{fig:optimal-irrigation-beta}).
As expected, a high $\beta$ leads to less irrigation at high water levels.
That's because citizens might be able to benefit from the remaining water again in future periods, while when using water for irrigation farmers can only benefit once in the present period.
Increasing $\beta$ means putting more weight on future utility up to a factor of $\beta = 1$, where stakeholders are indifferent between current and future utility.
A low $\beta$ on the other hand leads to a higher amount of water used for irrigation, especially at higher water levels. 
This can be easily imagined with the extreme case of $\beta = 0$, which means that no one benefits from leaving water in the reservoir for the next period.
Of course there is still water that's being left in the reservoir, but that is only to benefit citizens in the current period and not because of the benefits of the remaining water for the next period.
In this case the model is reduced to a single-stage optimization problem with no uncertainty, where the only task is the optimal allocation of the given current amount of water on the basis of the stakeholders' utility functions.
\\
We then went on and looked at the effect of increased rainfall volatility (\ref{fig:aggregated-use-rainfall-default} and \ref{fig:aggregated-use-rainfall-more-volatile}).
Because of the constant mean, the total amount of rain and therefore of water influx in the reservoir stays the same.
The result is a lower average water level in the reservoir.
The aggregated use is a good indicator of the efficiency our allocation strategy is achieving.
We could see that with rainfalls getting more volatile the sum of aggregated use over all simulated periods drops significantly, due to utility losses caused by increased likelihood of reservoir overflow and drought.
This vividly illustrates a very real consequence of climate change that governments and counties might face in the future.
\\
Instead of simulated rainfall, we also used real weather data from Barisal, Bangladesh. 
The real rainfall distribution (\ref{fig:rainfall-real-data}) looked quite a bit different then our lognormal distribution with $\mu=0$ and $\sigma=1$.
This indicates that other parameters will have to be found to better describe the annual rainfall in that specific area.
The higher expected value of rain in the real distribution leads to the optimal irrigation policy showing a higher amount of water used for irrigation at higher water levels (\ref{fig:optimalIrrigationPolicy-real-data}).
Further adding evaporation to the model lowers the average amount of water in the reservoir and therefore also the average amount of water used for irrigation and the sum of utility the optimal irrigation strategy gives us (\ref{fig:total-utility-evap}).
In the scenario with the highest scale factor for evaporation the reservoir runs completely dry in multiple periods, which leads to a utility of negative infinity in all of those periods.
This could be sensible in a scenario of an isolated village, where the survival of villagers completely depends on the local farmers' food production.
In this case our model would predict a famine, which under all circumstances has to be prevented.
An option would be to supply the village from the outside and assign a value to using this option to respect it in the model.
The value should somehow correlate to the effort this would take.
\\
Summing up, we created a basic model of a water allocation problem that can be solved to optimality by using dynamic programming and can thereby support decisionmakers for the allocation of limited water resources.
The model can work with random distributions and with real datasets and could of course also work with random distributions fitted to real datasets.
To get more satisfactory recommendations from the model additional parameters and/or policies could be introduced.
For example it would make sense to consider the temperature during the summer months when it comes to determine the utility citizens get from recreational water use.
An additional policy that might be worth considering in the context could be food supply from external sources, as suggested above.
\\
As we learned from our research, the discount factor plays a big role when it comes to finding the optimal irrigation strategy.
A whole different problem, more of an empirical nature however, would therefore be the determination of the "correct" discount factor, whatever that may be.
This is a different area of research and a lot of publications are dedicated to this topic but further advancements could greatly strengthen the theoretical foundation of our model.

\clearpage
\printbibliography

\newpage

%----------------------------------------------------------------------------------------
%	STATUTORY DECLARATION IN LIEU OF AN OATH
%----------------------------------------------------------------------------------------
\pagestyle{empty} 
\begingroup
\begin{center}
	\large
	\textbf{Eidesstattliche Versicherung}
\end{center}

\vspace{0.3cm}

\begin{tabular}{@{}p{8cm}p{5.8cm}}
	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
	\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
	&\\
	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
	\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
	&\\
	\underline{\hspace{6cm}} & \underline{\hspace{5.8cm}} \\
	\vspace{0.02cm}Name, Vorname & \vspace{0.02cm}Matrikelnummer \\
\end{tabular}
\vspace{0.3cm}

Ich versichere hiermit an Eides Statt, dass ich die vorliegende Arbeit mit dem Titel

\begin{center}
	{\makeatletter{\emph{{\@title}}}\makeatother}
\end{center}

selbständig und ohne unzulässige fremde Hilfe erbracht habe. Ich habe keine anderen als die angegebenen Quellen und Hilfsmittel benutzt. Für den Fall, dass die Arbeit zusätzlich auf einem Datenträger eingereicht wird, erkläre ich, dass die schriftliche und die elektronische Form vollständig übereinstimmen. Die Arbeit hat in gleicher oder ähnlicher Form noch keiner Prüfungsbehörde vorgelegen.

\vspace{0.6cm}

\begin{tabular}{@{}p{8cm}p{5.8cm}}
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
	&\\
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
	&\\
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
\end{tabular}

\vspace{0.6cm}

\begin{small}
	\textbf{Belehrung:}\\
	Wer vorsätzlich gegen eine die Täuschung über Prüfungsleistungen betreffende Regelung einer Hochschulprüfungsordnung verstößt, handelt ordnungswidrig. Die Ordnungswidrigkeit kann mit einer Geldbuße bis zu 50 000 Euro geahndet werden. Zuständige Verwaltungsbehörde für die Verfolgung und Ahndung von Ordnungswidrigkeiten ist die Kanzlerin oder der Kanzler der RWTH Aachen. Im Falle eines mehrfachen oder sonstigen schwerwiegenden Täuschungsversuches kann der Prüfling zudem exmatrikuliert werden (§ 63 Abs. 5 HG NRW).

	\textbf{§ 156 StGB: Falsche Versicherung an Eides Statt}\\
	Wer vor einer zur Abnahme einer Versicherung an Eides Statt zuständigen Behörde eine solche Versicherung falsch abgibt oder unter Berufung auf eine solche Versicherung falsch aussagt, wird mit Freiheitsstrafe bis zu drei Jahren oder mit Geldstrafe bestraft.

	\textbf{§ 161 StGB: Fahrlässiger Falscheid; fahrlässige falsche Versicherung an Eides Statt}\\
	(1) Wenn eine der in den §§ 154 bis 156 bezeichneten Handlungen aus Fahrlässigkeit begangen worden ist, so tritt Freiheitsstrafe bis zu einem Jahr oder Geldstrafe ein.\\
	(2) Straflosigkeit tritt ein, wenn der Täter die falsche Angabe rechtzeitig berichtigt. Die Vorschriften des § 158 Abs. 2 und 3 gelten entsprechend
\end{small}

\vspace{0.6cm}

Die vorstehende Belehrung habe ich zur Kenntnis genommen:

\vspace{0.6cm}
\begin{tabular}{@{}p{8cm}p{5.8cm}}
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
	&\\
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
	&\\
	\underline{\smash{Aachen, den \hspace{4cm}}} & \underline{\hspace{5.8cm}}\\
	Ort, Datum & Unterschrift \\
\end{tabular}

\endgroup
\clearpage

\end{document}